---
title: "Getting and Cleaning Data Course Week 4 Project"
author: "Toby Jones"
date: "August 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## preperation.R

This section connects to the website, downloads the zip file, and unzips it.

```{r, eval=FALSE}
    if(!file.exists("./data")){dir.create("./data")}
    fileUrl <- "http://archive.ics.uci.edu/ml/machine-learning-databases/00240/UCI%20HAR%20Dataset.zip"
    download.file(fileUrl, destfile = "./data/UCI%20HAR%20Dataset.zip")
    unzip('./data/UCI%20HAR%20Dataset.zip', exdir = "./data")
```


This section reads in the activity labels and the features.
```{r, eval=FALSE}
    temp = read.table(paste(dataDir, "activity_labels.txt", sep = ""), sep = "")
    activityLabels = as.character(temp$V2)
    temp = read.table(paste(dataDir, "features.txt", sep = ""), sep = "")
    attributeNames = temp$V2
```

This section performs the following processing for the training data:

* reads in the training data
* Convert all columns to numeric
* Sets the column names
* Reads in the Activity and sets it as a factor
* Reads in the subjects and sets it as a factor
* Combines the training set, Activity, and subjects into one data frame
```{r, eval=FALSE}
    training_x <- read.table(paste(dataDir, "train/X_train.txt", sep = ""), sep = "")
    training_x[,names(training_x)] <- apply(training_x[,names(training_x)], 2, function(x) as.numeric(as.character(x)))
    names(training_x) = attributeNames
    training_y = read.table(paste(dataDir, "train/y_train.txt", sep = ""), sep = "")
    names(training_y) = "Activity"
    training_y$Activity = as.factor(training_y$Activity)
    levels(training_y$Activity) = activityLabels
    training_subjects = read.table(paste(dataDir, "train/subject_train.txt", sep = ""), sep = "")
    names(training_subjects) = "subject"
    training_subjects$subject = as.factor(training_subjects$subject)
    train = cbind(training_x, training_subjects, training_y)
```

This section performs the following processing for the test data:

* reads in the training data
* Convert all columns to numeric
* Sets the column names
* Reads in the Activity and sets it as a factor
* Reads in the subjects and sets it as a factor
* Combines the training set, Activity, and subjects into one data frame
```{r, eval=FALSE}
    testing_x = read.table(paste(dataDir, "test/X_test.txt", sep = ""), sep = "")
    testing_x[,names(testing_x)] <- apply(testing_x[,names(testing_x)], 2, function(x) as.numeric(as.character(x)))
    names(testing_x) = attributeNames
    testing_y = read.table(paste(dataDir, "test/y_test.txt", sep = ""), sep = "")
    names(testing_y) = "Activity"
    testing_y$Activity = as.factor(testing_y$Activity)
    levels(testing_y$Activity) = activityLabels
    testing_subjects = read.table(paste(dataDir, "test/subject_test.txt", sep = ""), sep = "")
    names(testing_subjects) = "subject"
    testing_subjects$subject = as.factor(testing_subjects$subject)
    test = cbind(testing_x, testing_subjects, testing_y)
```

This section saves out the training and test sets into one file and removes all of the variables created.
```{r, eval=FALSE}
    save(train, test, file = dataFile)
    rm(train, test, temp, training_y, testing_y, training_x, testing_x, 
       training_subjects, testing_subjects, activityLabels, attributeNames,
       dataDir, dataFile, fileUrl)
```

## run_analysis.R

This section loads the datafile created in the preperation.R script and lists the data frames loaded from the datafile.
```{r, eval=FALSE}
    dataFile <- './data/dataset.RData'
    load(dataFile)
    ls()
```

This section merges the training and test sets.
```{r, eval=FALSE}
    # 1. Merges the training and the test sets to create one data set.
    mergeData <- rbind(test,train)
```

This section extracts the mean and standard deviation.
```{r, eval=FALSE}
    # 2. Extracts only the measurements on the mean and standard deviation for each measurement.
    subMergeData <- mergeData[,-which(names(mergeData) %in% c("Activity", "subject"))]
    means <- colMeans(subMergeData, na.rm = TRUE)
    std_devs <- apply(subMergeData, 2, sd)
```

This section merges the training and test sets.
```{r, eval=FALSE}
    # 1. Merges the training and the test sets to create one data set.
    mergeData <- rbind(test,train)
```

This section shows the activity counts that were added in the preperation file.
```{r, eval=FALSE}
    # 3. Uses descriptive activity names to name the activities in the data set
    ## Added in preperation.R
    table(mergeData[,'Activity'])
```

This section shows the column names that were added in the preperation file.
```{r, eval=FALSE}
    # 4. Appropriately labels the data set with descriptive variable names.
    ## Added in preperation.R
    colnames(mergeData)
```

This section creates and saves a file of the data in Tidy form.
```{r, eval=FALSE}
    # 5. From the data set in step 4, creates a second, independent tidy data set 
    #    with the average of each variable for each activity and each subject.
    TidyData <- aggregate(. ~ subject+Activity, mergeData, mean)
    write.table(TidyData, file="./data/Tidy.txt", row.names = FALSE)
    rm(dataFile, test, train, mergeData, subMergeData, means, std_devs, TidyData)
```
